GEE models

### 2017-07-01 doc/manuscript.Rmd
moved methods for plsda/lcmm to vig, updated inline code

While GEE accounts for the longitudinal design of the data, this approach is
limited in that it cannot analyze the inherent multivariate nature of the
composition of the NEFA fraction. Therefore, to confirm the results of the GEE
analyses in a multivariate environment (all NEFA variables in a single model),
partial least squares discriminant analysis (PLS-DA) was used to identify the
patterns of NEFA composition against beta-cell function (using ISSI-2, which had
the majority of the associations in the GEE models) and for those
who converted to or maintained dysglycemia status (either IFG, IGT, or DM) over
the 6 years. Latent class mixed models (LCMM) were used to extract the underlying
trajectories of beta-cell function (ISSI-2) over the 6 years. For more detailed
explanation of GEE, PLS-DA, and LCMM, please see the ESM Methods.

All analyses were performed using R `r getRversion()` [@Rbase], along with the R
packages geepack `r packageVersion('geepack')` for GEE [@Hoejsgaard2006a], caret
`r packageVersion('caret')` for PLS-DA, and lcmm `r packageVersion('lcmm')` for
LCMM. The R code for this manuscript is available at 
https://dx.doi.org/10.6084/m9.figshare.3472433
. Results were considered statistically significance at p<0.05,
after adjusting for multiple testing using the Benjamini-Hochberg False
Discovery Rate [@Benjamini1995a]. STROBE was used as a guideline for reporting [@Vandenbroucke2007b].

@@

```{r variableAndDataSet, message=FALSE, results='hide'}
gee_df <- get_gee_data(ds)
dysgly_data <- get_dysglycemia_data(ds)

outcomes_is <- c('linvHOMA', 'lISI')
outcomes_bcf <- c('lIGIIR', 'lISSI2')
outcomes <- c(outcomes_is, outcomes_bcf)
ne_pct <- grep('^pct_ne', names(gee_df), value = TRUE)
ne_conc <- c('TotalNE', grep('^ne\\d\\d', names(gee_df), value = TRUE))
covariates <- c('VN', 'Waist', 'Sex', 'Ethnicity', 'BaseAge', 'ALT', 'FamHistDiab')
```

```{r geeData, cache=TRUE}
gee_results <- dplyr::bind_rows(
    analyze_gee(gee_df, outcomes_is, ne_pct, covariates, 'mol%', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_is, ne_conc, covariates, 'nmol/mL', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_bcf, ne_pct, covariates, 'mol%', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_bcf, ne_conc, covariates, 'nmol/mL', adj.p = TRUE)
    ) %>% 
    tidy_gee_results()

gee_results_unadj <- dplyr::bind_rows(
    analyze_gee(gee_df, outcomes_is, ne_pct, 'VN', 'mol%', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_is, ne_conc, 'VN', 'nmol/mL', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_bcf, ne_pct, 'VN', 'mol%', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_bcf, ne_conc, 'VN', 'nmol/mL', adj.p = TRUE)
    ) %>% 
    tidy_gee_results()
```

```{r lcmmPls, cache=TRUE, results='hide'}
# include also the 3-yr and 6-yr data for training/testing.
lcmm_results <- analyze_lcmm(ds)
lcmm_data <- combine_lcmm_data(ds, lcmm_results)
plsda_results <- analyze_plsda(lcmm_data)
plsda_dysgly <- dysgly_data %>% 
    dplyr::mutate(ConvertDysgly = factor(ConvertDysgly, levels = c(0, 1), labels = c('No', 'Yes'))) %>% 
    analyze_plsda('ConvertDysgly')
```

```{r inlineResults}
percent.nefa <- calculate_percent_nefa_contribution(ds) 
misclass.n.percent.bcf <- calculate_plsda_misclass(plsda_results)
misclass.n.percent.dysgly <- calculate_plsda_misclass(plsda_dysgly)
percent.change.outcomes <- calculate_outcomes_pct_change(ds)
lcmm.groups <- calculate_ngroups_lcmm(lcmm_results)
dm.convert <- calculate_conversion_dysgly(dysgly_data, 'ConvertDM')
predm.convert <- calculate_conversion_dysgly(dysgly_data, 'ConvertPreDM')
sex.npct <- calculate_factor_npercent(ds0yr$Sex, 'Female')
ethn.npct <- calculate_factor_npercent(ds0yr$Ethnicity, 'European')
fhd.npct <- calculate_factor_npercent(ds0yr$FamHistDiab, 'Yes')
est.gee <- calculate_gee_magnitude(gee_results)
sample.outcomes <- calculate_sample(ds)
```

>>>>

While GEE accounts for the longitudinal design of the data, this approach is
limited in that it cannot analyze the inherent multivariate nature of the
composition of the NEFA fraction. Therefore, to confirm the results of the GEE
analyses in a multivariate environment (all NEFA variables in a single model),
partial least squares

All analyses were performed using R `r getRversion()` [@Rbase], along with the R
packages geepack `r packageVersion('geepack')` for GEE [@Hoejsgaard2006a].
The R code for this manuscript is available at 
https://dx.doi.org/10.6084/m9.figshare.3472433. Results were considered statistically significance at p<0.05,
after adjusting for multiple testing using the Benjamini-Hochberg False
Discovery Rate [@Benjamini1995a]. STROBE was used as a guideline for reporting [@Vandenbroucke2007b].

@@

```{r inline_results}
pct_nefa <- calc_pct_nefa(project_data)
chg_outcomes <- calc_outcomes_chg_n(project_data)
dm_incid <- calc_conversion_dysgly('ConvertDM')
predm_incid <- calc_conversion_dysgly('ConvertPreDM')

ds0yr <- dplyr::filter(project_data, VN == 0)
female <- calc_npercent(ds0yr$Sex, 'Female')
european <- calc_npercent(ds0yr$Ethnicity, 'European')
famhist <- calc_npercent(ds0yr$FamHistDiab, 'Yes')
est_gee <- calculate_gee_magnitude(gee_results)
```
